The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-11 GRB111117A.tex
Addresses Johans comments
Closes #14

\abstract{
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
their hosts to measure their redshifts. These measurements, however, are
critical to constrain sGRB progenitors and delay time models. Here, we present
spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
= 2.211$. This makes GRB~111117A the most distant high-confidence short duration GRB
detected to date. Our spectroscopic redshift supersedes a lower redshift value
for this burst previously estimated from photometry.

@@

We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
with the distributions of complete samples of both long and short GRBs. The lGRB
sample is from \citet{Arcodia2016} and the sGRB sample is from
\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have redshifts, and
likewise for 5 out of 16 in the short sample. Bursts without redshifts have been
excluded for both groups. GRB~111117A occupies a unique position in
Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
short sample. The short sample, excluding GRB~111117A, is located at low
redshifts ($z < 1$) and is found to populate a similar column density
environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.

@@

A single sGRB at high redshift does little in terms of constraining the redshift
distribution of sGRBs. In particular, other sGRB hosts could be missed because
they are intrinsically fainter and thus this high-$z$ event is only detected due
to the brightness of it's host. \citet{Berger2014} compiled a sample of sGRB
host luminosities, normalized by the characteristic galaxy luminosity at their
respective redshift, $L_B/L^{\star}_{B}$. Out of the 39 hosts in the sample, 26
(66 per cent) have redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
the Schechter function for blue galaxies ($U - V < 0.25$) in the redshift window
$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007}. For GRB~111117A, we obtain
$L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in
\citet{Berger2014} with measured $L_B/L^{\star}_{B}$.

@@

In this \emph{Letter}, we have provided a revised, spectroscopic redshift for
the short GRB~111117A based on emission lines setting it at $z = 2.211$. This
value supersedes the previous photometric redshift of \citet{Margutti2012} and
\citet{Sakamoto2013}. %The erroneous redshift estimate of previous authors is
%attributed to a discrepancy in the measured $z$-band magnitude.

>>>>

\abstract{
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
their hosts to measure their redshifts. These measurements, however, are
critical to constrain sGRB progenitors and delay time models. Here, we present
spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
= 2.211$. This makes GRB~111117A the most distant high-confidence short duration
GRB detected to date. Our spectroscopic redshift supersedes a lower photometric
redshift value for this burst previously estimated.

@@

We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
with the distributions of complete samples of both long and short GRBs. The lGRB
sample is from \citet{Arcodia2016} and the sGRB sample is from
\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have measured redshifts, and
likewise for 5 out of 16 in the short sample. Bursts without measured redshifts have been
excluded for both groups. GRB~111117A occupies a unique position in
Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
short sample. The short sample, excluding GRB~111117A, is located at low
redshifts ($z < 1$) and is found to populate a similar column density
environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column density for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.

@@

A single sGRB at high redshift does little in terms of constraining the redshift
distribution of sGRBs. In particular, other sGRB hosts could be missed because
they are intrinsically fainter and thus this high-$z$ event is only detected due
to the brightness of it's host. \citet{Berger2014} compiled a sample of sGRB
host luminosities, normalized by the characteristic galaxy luminosity at their
respective redshift, $L_B/L^{\star}_{B}$. Out of the 39 hosts in the sample, 26
(66 per cent) have measured redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
the Schechter function for blue galaxies ($U - V < 0.25$) in the redshift window
$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007}. For GRB~111117A, we obtain
$L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in
\citet{Berger2014} with measured $L_B/L^{\star}_{B}$.

@@

In this \emph{Letter}, we have provided a revised, spectroscopic redshift
measurement for the short GRB~111117A based on host galaxy emission lines
setting it at $z = 2.211$. This value supersedes the previous photometric
redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. 
%The erroneous redshift estimate of previous authors is attributed to a
%discrepancy in the measured $z$-band magnitude.
